#Create the data frame from XML file

#load the data frame called LatLon from data.rda. 

```{r}
load('data.rda')
dim(LatLon)
names(LatLon)
```

```{r}
#install.packages('XML')
library(XML)
```

```{r}
CIA_xml_parse <- xmlParse("factbook.xml.gz")
```

#Use XPath to extract the infant mortality and the CIA country codes from the XML tree

```{r}
CIA_code_root <- xmlRoot(x=CIA_xml_parse)
CIA_rootNode <- xmlRoot(x=CIA_code_root)
CIA_Infant_Mortality_Rate <- xpathSApply(doc=CIA_rootNode,path='//field[@name="Infant mortality rate"]')
#CIA_Infant_Mortality_Rate
```

```{r}
CIA_Infant_Mortality_Rate_Size <- xmlSize(CIA_Infant_Mortality_Rate[[1]])

Infant_Mortality <- c()
for(i in 1:CIA_Infant_Mortality_Rate_Size) {
  Infant_Mortality[i] = xmlGetAttr(node=CIA_Infant_Mortality_Rate[[1]][[i]], name="number") 
  Infant_Mortality = as.numeric(Infant_Mortality[!is.na(Infant_Mortality)])
}
Infant_Mortality
```

```{r}
Country_Code <- c()
for(i in 1:CIA_Infant_Mortality_Rate_Size) {
  Country_Code[i] = xmlGetAttr(node=CIA_Infant_Mortality_Rate[[1]][[i]], name="country")
  Country_Code = toupper(Country_Code[!is.na(Country_Code)])
}
Country_Code
```

#Create a data frame called IM using this XML file. The data frame should have 2 columns: for Infant Mortality and CIA.Codes.

```{r}
IM <- data.frame("Infant Mortality" = Infant_Mortality, "CIA.Codes" = Country_Code)
head(IM)
```

#Extract the country populations from the same XML document. Create a data frame called Pop using these data.This data frame should also have 2 columns, for Population and CIA.Codes.

```{r}
CIA_population <- xpathApply(doc = CIA_rootNode,path='//field[@name="Population"]')
CIA_Population_Size <- xmlSize(CIA_population[[1]]) 

Population <- c()
for(i in 1:CIA_Population_Size) {
  Population[i] = xmlGetAttr(node=CIA_population[[1]][[i]], name="number")
  Population = as.numeric(Population[!is.na(Population)])
}
Population
```

```{r}
Country_Code <- c()
for(i in 1:CIA_Population_Size) {
  Country_Code[i] = xmlGetAttr(node=CIA_population[[1]][[i]], name="country")
  Country_Code = toupper(Country_Code[!is.na(Country_Code)])
}
Country_Code
```

```{r}
Pop <- data.frame("Population" = Population, "CIA.Codes" = Country_Code)
head(Pop)
```
#Merge the two data frames to create a data frame called IMPop with 3 columns: IM, Pop, and CIA.Codes

```{r}
IMPop <- merge(x=IM, y=Pop)
head(IMPop)
```

#Now merge IMPop with LatLon to create a data frame called AllData that has 6 columns for Latitude, Longitude, CIA.Codes, Country Name, Population, and Infant Mortality 

```{r}
AllData <- merge(x=IMPop, y=LatLon)
head(AllData)
```

#Create a KML document for google earth visualization. 

```{r}
makeBaseDocument = function(){
### This code creates the template for KML document 
### Your code here
  xmlDoc = newXMLDoc()
  xmlRoot = newXMLNode("kml", doc = xmlDoc)
  xmlDocumentNode = newXMLNode("Document", parent = xmlRoot)
  newXMLNode(name="name", "Country Facts", parent = xmlDocumentNode)
  newXMLNode(name="description", "Infant mortality", parent = xmlDocumentNode)
  xmlLookAtNode = newXMLNode("LookAt", parent = xmlDocumentNode)
  newNode = newXMLNode(name="Folder",parent = xmlDocumentNode)
  newXMLNode(name = "Name","CIA Fact Book",parent = newNode)
  newXMLNode(name="longitude", "-121", parent = xmlLookAtNode)
  newXMLNode(name="latitude", "43", parent = xmlLookAtNode)
  newXMLNode(name="altitude", "4100000", parent = xmlLookAtNode)
  newXMLNode(name="title", "0", parent = xmlLookAtNode)
  newXMLNode(name="heading", "0", parent = xmlLookAtNode)
  newXMLNode(name="altitudeMode", "absolute", parent = xmlLookAtNode)
  return(xmlDoc)
}
```

```{r}
KmlDoc <- makeBaseDocument()
KmlRoot <- xmlRoot(KmlDoc)
KmlChildren <- xmlChildren(KmlDoc)
xmlDocumentNode <- KmlChildren[[1]][[1]]
KmlLookAtNode <- KmlChildren[[1]][[2]]
KmlDoc
```

```{r}
addPlacemark = function(lon, lat, ctryCode, ctryName, pop, infM, parent, inf1, pop1, style = FALSE) {
  pm = newXMLNode("Placemark", newXMLNode("name", ctryName), attrs = c(id = ctryCode),parent = parent)
  newXMLNode("description", paste(ctryName, "\n Population: ", pop,"\n Infant Mortality: ", infM, sep =""),parent = pm)

  point=newXMLNode("Point", newXMLNode("coordinates", paste(lon, ", ", lat, ", ", 0, sep="")), parent=pm)

  if(style) newXMLNode("styleUrl", paste("#YOR", inf1, "-", pop1, sep = ''), parent = pm)
}

for(i in 1:(dim(AllData)[1])) {
  addPlacemark(lon=AllData$Longitude[[i]],lat=AllData$Latitude[[i]],ctryCode=AllData$CIA.Codes[[i]], ctryName=AllData$Country.Name[[i]], 
    pop=AllData$Population[[i]], infM=AllData$Infant.Mortality[[i]], parent=xmlDocumentNode)
}

KmlDoc

```

#Use the two functions that to create the KML document and save it as 'Part2.kml'. open it in Google Earth. It #should have pushpins for all the countries.  


```{r}
saveXML(doc=KmlDoc, file="Part2.kml")
```

#Add Style to the KML

#Start fresh with a new KML document, by calling makeBaseDocument()

```{r}
KmlDoc2 = makeBaseDocument()
KmlRoot2 <- xmlRoot(KmlDoc2)
KmlChildren2 <- xmlChildren(KmlDoc2)
xmlDocumentNode2 <- KmlChildren2[[1]][[1]]
kmlLookAtNode2 <- KmlChildren2[[1]][[2]]

KmlDoc2
```
```{r}
infCut = cut(AllData$Infant.Mortality, breaks = c(0, 10, 25, 50, 75, 200))
infCut = as.numeric(infCut)
popCut = cut(log(AllData$Population), breaks = 5)
popCut = as.numeric(popCut)
```

#add styles and placemarks to KmlDoc2.

```{r}
scale = c(0.2, 1, 2.5, 3.5, 4.5)
colors = c("blue","green","yellow","orange","red")
```

```{r}
addStyle = function(col1, pop1, parent, DirBase, scales = scale)
{
  new_Col = colors[col1]
  st = newXMLNode("Style", attrs = c("id" = paste("YOR", col1, "-", pop1, sep="")), parent = parent)
  newXMLNode("IconStyle", 
             newXMLNode("scale", scales[pop1]), 
             newXMLNode("Icon", paste(DirBase, "color_label_circle_", colors[col1], ".png", sep ="")), parent = st)
}

root2 = xmlRoot(KmlDoc2)
DocNode = root2[["Document"]]
```

```{r}
for (k in 1:5)
{
  for (j in 1:5)
  {
    addStyle(j, k, DocNode, '/Users/aarthi/Downloads/color_label_circle/',scale)
  }
}


for(i in 1:(dim(AllData)[1])) {
  addPlacemark(lon=AllData$Longitude[[i]],lat=AllData$Latitude[[i]], ctryCode=AllData$CIA.Codes[[i]], ctryName=AllData$Country.Name[[i]],pop=AllData$Population[[i]], infM=AllData$Infant.Mortality[[i]],parent=xmlDocumentNode2, inf1=infCut[[i]], pop1=popCut[[i]], style=TRUE)
}
#KmlDoc2
```

#Finally, save your KML document, call it output.kml and open it in Google Earth to verify that it works.  

```{r}
saveXML(doc=KmlDoc2, file="output.kml")
```
